Order Reduction and MILP Models for Real-Time DNN Robustness Certification \\ Supplementary Material

### 2025-08-05 AAAI2026/PCA.tex
typos

PCA (Chinesta et al. 2017) is a linear technique to reduce the dimension of a dataset while keeping its main information.
From a dataset, the eigen vectors are computed, depicting the most important linear components of this dataset.
projecting over the first few main eigen vectors is a powerful model order reduction technique ({\em encoding} / {\em reducing from the full dimension}). 
It is easy to go from the reduce dimension back to the full dimension
by adding each of the first few eigen vectors times the coefficient in the reduced basis ({\em decoding}).

Figure~\ref{fig.MNIST} depicts the pipeline used for computing:

@@

\item {\em In exploitation}, we check whether a test dataset has the same accuracy (class predicted matches the ground truth) 
going directly through the MNIST DNN than when being encoded and decoded by the PCA decoding first.
The number of retained PCA dimensions (20) is selected to ensure that the MNIST DNN's classification accuracy on reconstructed images (i.e., after PCA encoding and decoding) remains the same $97$\% as the MNIST DNN used directly on the full-dimensional MNIST images.
%
\item {\em In Bound Computation}, we consider as input the 20 ($\times 2$: input and perturbed input) reduced PCA dimensions, from which all other variables are fixed.
However, we can define 784 $\times 2$ linear variables corresponding to the original image $I$ and its perturbation $I'$ before PCA encoding, as the constraints between these and the input are linear.
We fix the $L_1$-perturbation $\varepsilon$ on the image $I$ and its deformation $I'$, rather than on the inputs, as its where it is meaningful.
Finally, the goal is to optimize the value of $o_6 -o_8+ o'_8 - o'-6$,  where $o_C$ is the output neuron of the DNN corresponding to class $C \in \{0,\ldots, 9\}$.
\end{enumerate}

@@

To predict the plastic strain (which is not directly observable), we thus consider the deformation (which is directly observable), reduce it using the PCA reduction to obtain the reduced deformation, call the surrogate model, obtain the associated predicted reduced strain, and decode the reduced strain, obtaining the strain of the pipe in the full 3000 dimensions. Similarly to the MNIST case, we consider as input the 10 $\times 2$ dimensions of reduced deformation, but also consider 3000 $\times 2$ linear variables corresponding to the deformation and its perturbation, with linear constraints between them and the inputs (provided by the PCA matrix), and set the $L_1$ perturbation $\varepsilon$ over the Deformation and its perturbation, on which things are physically meaningful.
The surrogate outputs 26 output dimensions ($\times 2$) of reduced strain, but again 
as we have the linear PCA decoder explicitly, we can define the optimization goal 
in terms of optimizing the difference over 10 specific points in the full geometry (where thigs are physically relevant), rather than on the reduced strain space.

>>>>

PCA (Chinesta et al. 2017) is a linear technique to reduce the dimension of a dataset while keeping its main information.
From a dataset, the eigen vectors are computed, depicting the most important linear components of this dataset.
Projecting over the first few main eigen vectors is a powerful model order reduction technique ({\em encoding} / {\em reducing from the full dimension}). 
It is easy to go from the reduce dimension back to the full dimension
by making the product between the reduced vector and the first few eigen vectors ({\em decoding}).

Figure~\ref{fig.MNIST} depicts the MNIST pipeline on reduced dimensions, computing:

@@

\item {\em In exploitation}, we check whether a test dataset has the same accuracy (class predicted matches the ground truth) 
going directly through the MNIST DNN than when being encoded and decoded by the PCA decoding first.
The number of retained PCA dimensions (20) is selected to ensure that the MNIST DNN's classification accuracy on reconstructed images (i.e., after PCA encoding and decoding) remains the same $97$\% as the MNIST DNN used directly on the full-dimensional MNIST images.
It ensures no loss of accuracy.
%
\item {\em In Bound Computation}, we consider as input the 20 ($\times 2$: input and perturbed input') reduced PCA dimensions, from which all other variables are fixed.
However, we can define 784 $\times 2$ linear variables corresponding to the original image $I$ and its perturbation $I'$ before PCA encoding, as the constraints between these and the input are linear.
We fix the $L_1$-perturbation $\varepsilon$ on the image $I$ and its deformation $I'$, rather than on the inputs, as that is where it is meaningful.
Finally, the goal is to optimize the value of $o_6 -o_8+ o'_8 - o'_6$,  where $o_C$ is the output neuron of the DNN corresponding to class $C \in \{0,\ldots, 9\}$.
\end{enumerate}

@@

To predict the plastic strain (which is not directly observable), we thus consider the deformation (which is directly observable), reduce it using the PCA reduction to obtain the reduced deformation, call the surrogate model, obtain the associated predicted reduced strain, and decode the reduced strain, obtaining the strain of the pipe in the full 3000 dimensions. Similarly to the MNIST case, we consider as input the 10 $\times 2$ dimensions of reduced deformation, but also consider 3000 $\times 2$ linear variables corresponding to the deformation and its perturbation, with linear constraints between them and the inputs (provided by the PCA encoder, which is given under the form of a matrix), and set the $L_1$ perturbation $\varepsilon$ over the Deformation and its perturbation, where it is physically meaningful.
The surrogate outputs 26 output dimensions ($\times 2$) of reduced strain, but again 
as we have the linear PCA decoder explicitly, we can define the optimization goal 
in terms of optimizing the difference over 10 specific points in the full geometry (where thigs are physically relevant), rather than on the reduced strain space.
